---
title: "3. Prognostic Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{3. Prognostic Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(E2E)
```

# Prognostic Models (Survival Analysis)

This track is dedicated to survival prediction tasks.

## 1. Initialization

First, initialize the prognostic modeling system.

```{r}
initialize_modeling_system_pro()
```

## 2. Training Single Models with `models_pro`

The `models_pro` function trains one or more standard survival models. For this demonstration, we'll run a subset.

```{r}
# Run a subset of available prognostic models
results_all_pro <- models_pro(train_pro, model = c("lasso_pro", "rsf_pro"))

# Print summary for Random Survival Forest
print_model_summary_pro("rsf_pro", results_all_pro$rsf_pro)
```

## 3. Ensemble Modeling

### Bagging (`bagging_pro`)

Builds a Bagging ensemble for survival models.

```{r}
# Create a Bagging ensemble with GBM as the base survival model
# n_estimators is reduced for faster execution.
bagging_gbm_pro_results <- bagging_pro(train_pro, base_model_name = "gbm_pro", n_estimators = 5, seed = 123)
print_model_summary_pro("Bagging (GBM)", bagging_gbm_pro_results)
```

### Stacking (`stacking_pro`)

Builds a Stacking ensemble for survival models.

```{r}
# Create a Stacking ensemble with GBM as the meta-model
stacking_gbm_pro_results <- stacking_pro(
  results_all_models = results_all_pro,
  data = train_pro,
  meta_model_name = "gbm_pro"
)
print_model_summary_pro("Stacking (GBM)", stacking_gbm_pro_results)
```

## 4. Applying Models to New Data (`apply_pro`)

Generate prognostic scores for a new dataset.

```{r}
# Apply the trained stacking model to the test set
pro_pred_new <- apply_pro(
  trained_model_object = stacking_gbm_pro_results$model_object,
  new_data = test_pro,
  time_unit = "day"
)

# Evaluate the new prognostic scores
eval_pro_new <- evaluate_model_pro(
  precomputed_score = pro_pred_new$score,
  Y_surv_obj = survival::Surv(time = test_pro$time, event = test_pro$outcome),
  sample_ids = test_pro$sample,
  years_to_evaluate = c(1, 3, 5)
)
print(eval_pro_new$evaluation_metrics)
```

## 5. Visualization (`figure_pro`)

Generate Kaplan-Meier (KM) and time-dependent ROC (tdROC) curves.

```{r, fig.width=6, fig.height=5, warning=FALSE}
# Kaplan-Meier Curve (don't save to file in vignette)
figure_pro(type = "km", data = stacking_gbm_pro_results, time_unit= "days", output_file = "Prognostic_KM",
           output_type = "pdf")

# Time-Dependent ROC Curve
figure_pro(type = "tdroc", data = stacking_gbm_pro_results, time_unit = "days", output_file = "Prognostic_TDROC", output_type = "pdf")
```
